---
title: "StormData_Analysis"
output: html_document
---
## Title 
The goal of this project is to explore the NOAA Storm Database and answer some important questions about severe weather events.

## Synopsis
People are more concerned about the influences of severe weahter events on populatoin health and economy. So, this project investigates two related questions: the first question is "which types of events are most harmful with respect to population health?" and the second is "which types of events have the greatest economic consequences?" In the following analysis, these two questions can be answered by some tables and clear plots. 

## Data processing
```{r,cache=TRUE}
storm_data <- read.csv(bzfile("repdata-data-StormData.csv.bz2"))
```
Event types don't have specific format, so we need to do some preprocessing.
```{r}
event_type <-  tolower(storm_data$EVTYPE)
event_type <- gsub("[[:blank:][:punct:]+]","",event_type)
storm_data$EVTYPE <- event_type
````

## Results
### Which types of events are most harmful with respect to population health?
```{r}
event_fatality <- aggregate(FATALITIES~EVTYPE,storm_data,sum)
event_injury <- aggregate(INJURIES~EVTYPE,storm_data,sum) 
```
List top 8 events that caused the most number of deaths 
```{r}
t1 <- head(event_fatality[order(event_fatality$FATALITIES,decreasing = T),],8);t1
```
List top 8 events that caused the most number of injuries 
```{r}
t2 <- head(event_injury[order(event_injury$INJURIES,decreasing = T),],8);t2
```

### Plots of different weather event types with respect to population health
```{r}
library(ggplot2)
library(cowplot)
g1 <- ggplot(t1,aes(x=reorder(EVTYPE,FATALITIES),y=FATALITIES))+
  geom_bar(stat="identity",fill="seagreen")+
  labs(x="Event Type",y="Total Number of Deaths",title="Severe Weather Events w.r.t Fatalities")+
  theme(axis.text=element_text(size=8))
g2 <- ggplot(t2,aes(x=reorder(EVTYPE,INJURIES),y=INJURIES))+
  geom_bar(stat="identity",fill="seagreen")+
  labs(x="Event Type",y="Total Number of Injuries",title="Severe Weather Events w.r.t Injuries")+
  theme(axis.text=element_text(size=8))
```
The following plots show severe weather events that resulted in deaths and injuries
```{r}
plot_grid(g1, g2,labels=c("A","B"), nrow = 2)
```

### Which types of events have the greatest economic consequences?

In the raw data, economic consequences are represented by two fields, which are PROPDMG and CROPDMG. These two fields can be used to analyze the property and crop damage for each event.
```{r}
# function for transforming letters to exponents
expo <- function(x){
  if (x %in% c("h", "H"))
        return(2)
    else if (x %in% c("k", "K"))
        return(3)
    else if (x %in% c("m", "M"))
        return(6)
    else if (x %in% c("b", "B"))
        return(9)
    else if (!is.na(as.numeric(x)))
        return(as.numeric(x))
    else if (x %in% c('', '-', '?', '+'))
        return(0)  
    else {
        stop("Invalid Value")
    }
} 
```
Transform economic losses into real values
```{r}
prop_dmg_exp <- sapply(storm_data$PROPDMGEXP,FUN=expo)
storm_data$PROPDMG <- storm_data$PROPDMG*(10**prop_dmg_exp) 
crop_dmg_exp <- sapply(storm_data$CROPDMGEXP,FUN=expo)
storm_data$CROPDMG <- storm_data$CROPDMG*(10**crop_dmg_exp) 
```
```{r}
event_prop <- aggregate(PROPDMG~EVTYPE,storm_data,sum)
event_crop <- aggregate(CROPDMG~EVTYPE,storm_data,sum)
```
List top 6 events that caused significant property damages
```{r}
t3 <- head(event_prop[order(event_prop$PROPDMG,decreasing = T),]);t3
```
List top 8 events that caused significant crop damages
```{r}
t4 <- head(event_crop[order(event_crop$CROPDMG,decreasing = T),],8);t4
```

### Plots of different weather event types with respect to economic impacts
```{r}
g3 <- ggplot(t3,aes(x=reorder(EVTYPE,PROPDMG),y=PROPDMG))+
  geom_bar(stat="identity",fill="seagreen")+
  labs(x="Event Type",y="Economic Losses",title="Property Damage caused by different weather events")+
  theme(axis.text=element_text(size=8))
g4 <- ggplot(t2,aes(x=reorder(EVTYPE,INJURIES),y=INJURIES))+
  geom_bar(stat="identity",fill="seagreen")+
  labs(x="Event Type",y="Economic Losses",title="Crop Damage caused by different weather events")+
  theme(axis.text=element_text(size=8))
```
The following plots show weather events that caused significant property damage and crop damage
```{r}
plot_grid(g3, g4,labels=c("A","B"), nrow = 2)
```




