---
Title: "statistical_inference_project_part1"
Author: Derek Chang
output: pdf_document
---
## Title: statistical_inference_project_part1
## Author: Derek Chang

## Overview
This is a project for statistical inference class on Coursera. There will two parts in this project: 1. A simulation exercise 2. Basic inferential data analysis. The first part is going to be reported on this PDF file.

## Simulation Exercise

### Show the sample mean and compare it to the theoretical mean of the distribution
```{r}
# load required package
library(ggplot2)
# set rate parameter to be 0.2
lambda = 0.2
# sample number 40
n = 40 
# simulation times 1000
simulation = 1000
set.seed(123)
# run the simulation 1000 times and compute the mean of each 40 exponentials
simulated_exponents = replicate(simulation,rexp(n,lambda))
averages = apply(simulated_exponents,2,mean)
# compute the average of the distribution of the mean of 40 exponentials
sample_mean = mean(averages);sample_mean
# compute theretical mean of the distribution
theoretical_mean = 1 / lambda;theoretical_mean 
# plot the distribution of the mean of 40 exponentials
p1 = ggplot(as.data.frame(averages),aes(x=averages))+
  geom_histogram(fill="cadetblue",colour="black")+
  geom_vline(xintercept = c(sample_mean,theoretical_mean),colour="darkgreen",linetype="longdash")+
labs(x="the mean of 40 exponentials",title="The distribution of the mean of 40 exponentials")
p1
```
In the plot, we can see that theoretical mean (5) and sample mean (5.011911) are very close to each other.

### Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution
```{r}
# sample variance and standard deviation
sample_variance = sd(averages)^2;sample_variance
sample_std = sqrt(sample_variance);
# theoretical variance
theoretical_variance = (1/lambda)^2/n;theoretical_variance
# plot for comparing the sample variance to the theoretical variance of the distribution
variance_data = data.frame(variances=c(theoretical_variance,sample_variance))
rownames(variance_data) = c("theorical variance","sample variance")
p2 = ggplot(variance_data,aes(x=rownames(variance_data),y=variances))+
  geom_bar(stat="identity",fill="cadetblue")+
  labs(x="",title="Comparison between sample and theoretical variances of the distribution")
p2
```
In this figure, sample variance (0.6004928) and theoretical variance (0.625) are very close.

### Show that the distribution is approximately normal
```{r}
p3 = ggplot(as.data.frame(as.numeric(simulated_exponents)),aes(x=averages))+
  geom_histogram(aes(y=..density..), fill="cadetblue",colour="black")+
  geom_density(colour="red",size=1)+
  labs(x="",title="The density of the distribution of random exponentials")
p3
qqnorm(simulated_exponents);qqline(simulated_exponents,col=2)
p4 = ggplot(as.data.frame(averages),aes(x=averages))+
  geom_histogram(aes(y=..density..), fill="cadetblue",colour="black")+
  geom_density(colour="red",size=1)+
  labs(x="the mean of 40 exponentials",title="The density of the distribution of the mean of 40 exponentials")
p4
qqnorm(averages);qqline(averages,col=2)
```
From qqplot and the density of the distribution, the distribution of the mean of 40 exponentials is more like a normal distribution than that of a large collection of random exponentials.

